Covariate and differential expression analysis of ROSMAP reprocessed counts Using Diagnosis as a covariate for TWAS

### 2019-09-17 code/ROSMAP_RegressDiagnosis_4_TWAS.Rmd
Full Store Implemented

# Store enrichment results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'ROSMAP_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ = File('ROSMAP_SVA_Diagnosis_Corrected.tsv', 
                  name = 'SVA and Diagnosis Regressed Data', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'SVA_Diagnosis_Regressed'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Store file:
activity <- Activity(
    'Fancy Processing',
    description='No seriously, really fancy processing',
    used=c('syn1906480', 'http://data_r_us.com/fancy/data.txt'),
    executed='syn1917825')
file <- File('/path/to/data/file.xyz', description='Fancy new data', parent=project)
file <- synStore(file, activity=activity)

# Function to change entity ACL (also creates new acls if it doesn't exist)
synChangeEntityACL <- function(id, new.racl){
  tryCatch({ 
    # If old acl exist update
    old.acl = synGetEntityACL(id)
    old.acl@resourceAccess = new.racl
    synUpdateEntityACL(old.acl)
  }, error = function(e){ 
    # If not create a new acl
    new.acl = AccessControlList(id = id, resourceAccess = new.racl)
    synCreateEntityACL(new.acl)
  })
}

# Create new resource acls
new.racl = synapser::synSetPermissions(principalId = '3319864', 
                          accessType = c('UPDATE', 'CHANGE_SETTINGS', 'DELETE', 'READ', 
                                         'DOWNLOAD', 'MODERATE', 'CREATE', 'CHANGE_PERMISSIONS')) %>%
  ResourceAccessList() %>%
  synapseClient::append(synSetPermissions(principalId = '273995', 
                                       accessType = c('UPDATE', 'CHANGE_SETTINGS', 'DELETE', 'READ', 
                                                      'DOWNLOAD', 'MODERATE', 'CREATE', 'CHANGE_PERMISSIONS'))) %>%
  synapseClient::append(synSetPermissions(principalId = '3344250', 
                                       accessType = c('READ', 'DOWNLOAD')))
# Get annotations
all.annotations = annotations(COUNT_OBJ)
# Code
CODE <- Folder(name = "ROSMAP Reprocessed RNASeq (Normalised) Diagnosis Removed", parentId = parentId)
CODE <- synStore(CODE)
old.acl = synGetEntityACL(CODE@properties$id)
synChangeEntityACL(CODE@properties$id, new.racl)
# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'ROSMAP_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('ROSMAP_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
all.annotations@annotations@stringAnnotations$normalizationStatus	= TRUE
all.annotations@annotations@stringAnnotations$normalizationType	= 'CQN'
all.annotations@annotations@stringAnnotations$fileFormat = 'tsv'
all.annotations@annotations@stringAnnotations$dataSubType = 'covariates'
annotations(COV_OBJ) = all.annotations
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName,
                   executed = thisFile, activityDescription = activityDescription)
synChangeEntityACL(COV_OBJ@properties$id, new.racl)
# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('ROSMAP_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
all.annotations@annotations@stringAnnotations$dataSubType = 'filteredCounts'
annotations(COUNT_OBJ) = all.annotations
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# Store filtered counts
NEW.COUNTS %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_eCounts.tsv', sep = '\t', row.names=F, quote=F)
ECOUNT_OBJ = File('ROSMAP_DLPFC_eCounts.tsv', name = 'Counts (estimated)', parentId = CODE$properties$id)
all.annotations@annotations@stringAnnotations$dataSubType = 'estimatedCounts'
annotations(ECOUNT_OBJ) = all.annotations
ECOUNT_OBJ = synStore(ECOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# Store logCPM
CQN.GENE_EXPRESSION$y %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('ROSMAP_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
all.annotations@annotations@stringAnnotations$dataSubType = 'filteredLCPM'
annotations(LCOUNT_OBJ) = all.annotations
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('ROSMAP_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
all.annotations@annotations@stringAnnotations$dataSubType = 'offset'
annotations(OFFSET_OBJ) = all.annotations
OFFSET_OBJ = synStore(OFFSET_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# Store design matrix
MODEL1 %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_DLPFC_Design.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('ROSMAP_DLPFC_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
all.annotations@annotations@stringAnnotations$dataSubType = 'designMatrix'
annotations(DM_OBJ) = all.annotations
DM_OBJ = synStore(DM_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
synChangeEntityACL(DM_OBJ@properties$id, new.racl)
# Store surrogate variables
SURR.VAR %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_DLPFC_SurVar.tsv', sep = '\t', row.names=F, quote=F)
SURR_OBJ = File('ROSMAP_DLPFC_SurVar.tsv', name = 'Surrogate Variables', parentId = CODE$properties$id)
all.annotations@annotations@stringAnnotations$dataSubType = 'surrogateVariables'
annotations(SURR_OBJ) = all.annotations
SURR_OBJ = synStore(SURR_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# Store residual gene expression for eQTL analysis
RESIDUAL.SVA.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_eQTLResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
eEXP_OBJ = File('ROSMAP_DLPFC_eQTLResidualExpression.tsv', 
                name = 'Normalised, covariates, surrogate and Diagnosisvariable adjusted residual expression (for eQTL)',
                parentId = CODE$properties$id)
all.annotations@annotations@stringAnnotations$dataSubType = 'residualGeneExpForeQTL'
annotations(eEXP_OBJ) = all.annotations
eEXP_OBJ = synStore(eEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# Store residual gene expression for network analysis
RESIDUAL.NET.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('ROSMAP_DLPFC_netResidualExpression.tsv', 
                name = 'Normalised, covariates and Diagnosis removed residual expression (for network analysis)', 
                parentId = CODE$properties$id)
all.annotations@annotations@stringAnnotations$dataSubType = 'residualGeneExpForNetAnlz'
annotations(nEXP_OBJ) = all.annotations
nEXP_OBJ = synStore(nEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Delete folder acl
synChangeEntityACL(CODE@properties$id, old.acl@resourceAccess)
stopCluster(cl)
```

@@

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')
synapseLogin()
knit2synapse::knitfile2synapse(file = "./ROSMAP_RegressDiagnosis_4_TWAS.Rmd",
                                 owner = "syn20798014")#,
                                 #entityName = 'ROSMAP Reprocessed RNASeq (Normalised) Diagnosis Regressed Using #Diagnosis as Covariate')
```

>>>>

# Store SVA results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'ROSMAP_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ = File('ROSMAP_SVA_Diagnosis_Corrected.tsv', 
                  name = 'Normalised, covariates, surrogate and Diagnosis variable adjusted residual expression (for eQTL)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'residualGeneExpForeQTL'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'ROSMAP_SVA_Diagnosis_Covariates.tsv', row.names = F, sep = '\t', quote=F)
ENRICH_OBJ = File('ROSMAP_SVA_Diagnosis_Covariates.tsv', 
                  name = 'Covariates', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'covariates'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_Counts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_Counts.tsv', 
                  name = 'Counts (filtered raw)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'filteredCounts'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store Estimated counts
NEW.COUNTS %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_eCounts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_Counts.tsv', 
                  name = 'Counts (estimated)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'estimatedCounts'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store logCPM
CQN.GENE_EXPRESSION$y %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_logCPM.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_logCPM.tsv', 
                  name = 'Counts (filtered logCPM)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'filteredLCPM'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_offset.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_offset.tsv', 
                  name = 'Gene length and GC content offset', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'offset'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store design matrix
MODEL1 %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_Design.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_Design.tsv', 
                  name = 'Design Matrix', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'designMatrix'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store surrogate variables
SURR.VAR %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_SurVar.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_SurVar.tsv', 
                  name = 'Surrogate Variables', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'surrogateVariables'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store residual gene expression for network analysis
RESIDUAL.NET.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_netResidualExpression.tsv', 
                  name = 'Normalised, covariates and Diagnosis removed residual expression (for network analysis)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'residualGeneExpForNetAnlz'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

stopCluster(cl)
```

@@

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')
synapseLogin()
knit2synapse::knitfile2synapse(file = "./ROSMAP_RegressDiagnosis_4_TWAS.Rmd",
                                 owner = "syn20798014")
```
